---
output:
 html_document:
  theme: yeti
---

<head>
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
</head>

<style>

code{
  background-color:#f2f2f2;
  border-radius: 25px;
}
 
span.co{
  color:#000080;
  font-weight: bold;
}
 
img{
  display: block;
  padding-left: 15px;
  padding-right: 15px;
  padding-top: 10px;
  padding-bottom: 10px;
}

h4{
  text-align: left;
  font-size: 20px;
}

p{
  text-align: left;
  font-size: 16px;
}

ul, ol{
  line-height: 27px;
  text-align: left;
  font-size: 16px;
  margin-left: 0px;
}
 
blockquote{
  font-size: 18px;
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding-left: 16px;
  padding-right: 16px;
}
 
.row{
  margin: auto;
}
 
table {
  border-collapse: collapse;
}

table, td, th {
  border: 1px solid black;
  padding: 5px;
  text-align: center;
  vertical-align: middle;
}
 
 /* Create two equal columns that floats next to each other */
.column {
  float: left;
  width: 50%;
  padding: 10px;
}

/* Clear floats after the columns */
.row:after {
  content: "";
  display: table;
  clear: both;
}

.roundBorder {
  border-radius: 25px;
  border: 5px solid #30288C;
  background: #D6EAF8;
  padding-left: 20px;
  padding-right: 20px;
  padding-top: 10px;
  padding-bottom: 10px;
}

.roundBorderBlack {
  border-radius: 25px;
  border: 10px solid #D3D3D3;
  padding-left: 20px;
  padding-right: 20px;
  padding-top: 10px;
  padding-bottom: 10px;
}

.roundBorderBlackEx {
  border-radius: 5px;
  border: 5px solid #D3D3D3;
  padding-left: 5px;
  padding-right: 5px;
  padding-top: 2px;
}

.roundBorderEx {
  border-radius: 3px;
  border: 5px solid #30288C;
  background: #D6EAF8;
  padding-left: 5px;
  padding-right: 5px;
  padding-top: 2px;
}

.tt {
    position: relative;
    display: inline-block;
    class: inline; 
    font-weight: bold;
    font-family: "Avenir";
    font-size: 18px;
    border-bottom: 1px black;
}

/* Tooltip text */
.tt .ttText {
    visibility: hidden;
    font-weight: normal;
    font-size: 18px;
    width: 200px;
    background-color: black;
    border: 1px solid black;
    color: white;
    text-align: left;
    padding: 5px;
    border-radius: 6px;
    position: absolute;
    z-index: 1;
}

/* Show the tooltip text when you mouse over the tooltip container */
.tt:hover .ttText {
    visibility: visible;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr) ; library(kableExtra)

options(knitr.table.format = "html")

knitr::opts_chunk$set(message = FALSE)


knitr::opts_chunk$set(warning = FALSE)

# set up ------------------------------------------------------------------

library(tidyverse)


# functions ---------------------------------------------------------------

# Function to add region to a plot (based on siteId field):

addRegion <- function(x){
  xState <- str_sub(x, start = -3, end = -2)
  case_when(
    xState == 'GA' ~ 'Atlanta',
    xState == 'CO' ~ 'Colorado',
    xState == 'FL' ~ 'Gainesville',
    xState == 'NC' ~ 'Raleigh',
    xState == 'PA' ~ 'Pittsburgh',
    xState == 'MA' ~ 'Springfield',
    xState %in% c('MD','DC', 'VA') ~ 'Washington, DC',
    TRUE ~ 'noData'
  )
}

# Outlier functions:


outlierBottom <- function(x){
  median(x) - 2.5*mad(x)
}

outlierTop <- function(x){
  median(x) + 2.5*mad(x)
}

outlierFilter <- function(df, var, x){
  df %>%
    filter(var > outlierBottom(x)) %>%
    filter(var < outlierTop(x))
}

outlierFilter <- function(df, x){
  # Column to analyze:
  colName <- enquo(x)
  # Get values vector:
  values <- df %>% 
    select(!!colName) %>%
    filter((!!colName) < 1000) %>%
    unlist(use.names = FALSE)
  # Filter:
  df %>%
    filter(!is.na((!!colName))) %>%
    filter(
      (!!colName) > median(values) - 2.5*mad(values)) %>%
    filter(
      (!!colName) < median(values) + 2.5*mad(values)
      )
}

# add data ----------------------------------------------------------------


# captures

captures <-
  suppressWarnings(
    read_csv('nocaReport/captureTable.csv') %>%
      setNames(str_replace_all(names(.), 'Capture', '')) %>%
      select(siteID, date, obs:bandNumber, mass:sex) %>%
      mutate(tl  = as.numeric(tl))
  ) %>%
   mutate(region = addRegion(siteID)) %>%
  filter(!region %in% c('Colorado', 'noData')) %>%
  mutate(region = factor(
    region,
    levels = c('Gainesville', 'Atlanta', 'Raleigh', 'Washington, DC', 'Pittsburgh', 'Springfield'))
    )

# NOCA captures:

noca <-
  captures %>%
  filter(spp == 'NOCA')

# Impervious:

imp <- 
  read_csv('nocaReport/impervious.csv') %>%
  filter(distance == 500,
         siteID %in% unique(noca$siteID)
         )

nocaImp <-
  left_join(
    noca, 
    imp %>% 
      select(-region),
    by = 'siteID'
  ) %>%
  filter(!is.na(distance))
```

<h1>Northern Cardinal data summary</h1>
<hr>
<br>
<h2>Some capture summary statistics</h2>

<div class = "row">
<div class = "column" style = "width:50%">
<ul>
<li>Captures by type</li>
<ul>
<li>Band: 
```{r nBanded, echo = FALSE, results = 'asis'}
noca %>%
  filter(enc == 'B') %>%
  nrow() %>%
  cat()
```
</li>
<li>Recapture: 
```{r nRecapped, echo = FALSE, results = 'asis'}
noca %>%
  filter(enc == 'R') %>%
  nrow() %>%
  cat()
```
</li>
</ul>
<li>Captures by age:</li>
<ul>
<li>Hatch year: 
```{r nHy, echo = FALSE, results = 'asis'}
noca %>%
  filter(age %in% c('HY', 'L')) %>%
  nrow() %>%
  cat()
```
</li>
<li>After hatch year: 
```{r nAHY, echo = FALSE, results = 'asis'}
noca %>%
  filter(age %in% c('AHY', 'ASY', 'SY')) %>%
  nrow() %>%
  cat()
```
</li>
<li>Unknown: 
```{r nUnknown_age, echo = FALSE, results = 'asis'}
noca %>%
  filter(age %in% c('noData', 'U')) %>%
  nrow() %>%
  cat()
```
</li>
</ul>
<li>Captures by sex:</li>
<ul>
<li>Female: 
```{r nFemale, echo = FALSE, results = 'asis'}
noca %>%
  filter(sex  == 'F') %>%
  nrow() %>%
  cat()
```
</li>
<li>Male: 
```{r nMale, echo = FALSE, results = 'asis'}
noca %>%
  filter(sex == 'M') %>%
  nrow() %>%
  cat()
```
</li>
<li>Unknown: 
```{r nUnknown_sex, echo = FALSE, results = 'asis'}
noca %>%
  filter(sex %in% c('noData', 'U')) %>%
  nrow() %>%
  cat()
```
</li>
</ul>
</ul>
</div>

<div class = "column" style = "width:50%">
<ul>
<li>Captures by region:</li>
<ul>
<li>Atlanta, GA: 
```{r regions, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Atlanta') %>%
  nrow() %>%
  cat()
```
</li>
<li>Gainesville, FL: 
```{r, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Gainesville') %>%
  nrow() %>%
  cat()
```
</li>
<li>Pittsburgh, PA: 
```{r, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Pittsburgh') %>%
  nrow() %>%
  cat()
```
</li>
<li>Raleigh, NC: 
```{r, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Raleigh') %>%
  nrow() %>%
  cat()
```
</li>
<li>Springfield, MA: 
```{r, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Springfield') %>%
  nrow() %>%
  cat()
```
</li>
<li>Washington, DC: 
```{r, echo = FALSE, results = 'asis'}
noca %>%
  filter(region == 'Washington, DC') %>%
  nrow() %>%
  cat()
```
</li>
</ul>
</ul>
</div>
</div>

<hr>
<h2>Some summary plots</h2>

```{r massPlot, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
noca %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(mass) %>%
  ggplot(aes(x = region, y = mass)) +
  geom_boxplot(aes(fill = region)) +
  facet_wrap(~sex) +
  coord_flip() +
   labs(title = 'Northern Cardinal mass by region',
       x = 'Mass (g)',
       y = 'Region') +
  scale_y_continuous(expand = c(0,0),
                     limits = c(30, 55)) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<br>
```{r wingPlot, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
noca %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(wing) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = region, y = wing)) +
  geom_boxplot(aes(fill = region)) +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal wing chord length by region',
       x = 'Wing chord (mm)',
       y = 'Region') +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(80, 100)) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<br>
```{r tailPlot, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
noca %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(tl) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = region, y = tl)) +
  geom_boxplot(aes(fill = region)) +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal tail length by region',
       x = 'Tail (mm)',
       y = 'Region') +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(80, 120)) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<br>
```{r massWing, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
noca %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(mass) %>%
  outlierFilter(wing) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = mass, y = wing)) +
  geom_point(aes(color = region), alpha = 0.2) +
  stat_smooth(aes(color = region, fill = region), method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal mass:wing by region',
       x = 'Mass (g)',
       y = 'Wing (mm)') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<hr>
<h2>Summarizing measurements with impervious</h2>
<br>
```{r densityImp, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  ggplot(aes(x = imp)) +
  geom_density(aes(fill = region)) +
  facet_wrap(~region) +
  theme_bw() +
  scale_x_continuous(expand = c(0,0),
                     limits = c(0, 100)) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, .07)) +
  labs(title = 'Density distribution of Northern Cardinal captures across the rural-to-urban gradient',
       x =  'Proportion of impervious surface within 500m',
       y = 'Point density') +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<br>
```{r massImpAcross, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(mass) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = imp, y = mass)) +
  geom_point(alpha = 0.2) +
  stat_smooth(method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal mass by impervious surface across regions',
       x = 'Proportion of impervious surface within 500m',
       y = 'Mass (g)') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```

<br>
```{r wingImpAcross, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(wing) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = imp, y = wing)) +
  geom_point(alpha = 0.2) +
  stat_smooth(method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal wing chord length by impervious surface across regions',
       x = 'Proportion of impervious surface within 500m',
       y = 'Wing chord length (mm)') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```

<br>
```{r massWingImpAcross, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(wing) %>%
  outlierFilter(mass) %>%
  mutate(
    sex = factor(sex, labels = c('Female', 'Male')),
    massWing = mass/wing) %>%
  ggplot(aes(x = imp, y = massWing)) +
  geom_point(alpha = 0.2) +
  stat_smooth(method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal mass:wing ratio by impervious surface across regions',
       x = 'Proportion of impervious surface within 500m',
       y = 'Mass:wing ratio') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
<br>

```{r massImp, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(mass) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = imp, y = mass)) +
  geom_point(aes(color = region), alpha = 0.2) +
  stat_smooth(aes(color = region, fill = region), method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal mass by impervious surface by region',
       x = 'Proportion of impervious surface within 500m',
       y = 'Mass (g)') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```

<br>
```{r wingImp, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(wing) %>%
  mutate(sex = factor(sex, labels = c('Female', 'Male'))) %>%
  ggplot(aes(x = imp, y = wing)) +
  geom_point(aes(color = region), alpha = 0.2) +
  stat_smooth(aes(color = region, fill = region), method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal wing chord length by impervious surface by region',
       x = 'Proportion of impervious surface within 500m',
       y = 'Wing chord length (mm)') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```

<br>
```{r massWingImp, echo = FALSE, results = 'asis', fig.width = 8, fig.asp = .5}
nocaImp %>%
  filter(sex %in% c('M', 'F')) %>%
  outlierFilter(wing) %>%
  outlierFilter(mass) %>%
  mutate(
    sex = factor(sex, labels = c('Female', 'Male')),
    massWing = mass/wing) %>%
  ggplot(aes(x = imp, y = massWing)) +
  geom_point(aes(color = region), alpha = 0.2) +
  stat_smooth(aes(color = region, fill = region), method = 'glm') +
  facet_wrap(~sex) +
  labs(title = 'Northern Cardinal mass:wing ratio by impervious surface by region',
       x = 'Proportion of impervious surface within 500m',
       y = 'Mass:wing ratio') +
  theme_bw() +
  theme(
    panel.spacing.x = unit(1, "lines")
  )
```
